Formal Verification of Real Models from Platforms Using The B-method

### 2009-04-21 trunk/papers/FM2009/source/paper.tex
just some changes

%\usepackage{float}

@@

\begin{abstract}
This paper describes an approach to model the microcontroller platforms. More specifically, it shows details about
the Z80 model. The model is created using the B method; it applies math and logic concepts to describe the
characteristics from platforms. Therefore, this model can be used in platforms projects to document, build
simulators, verify properties about the model and verify the software in assembly level. The verification in
assembly is utility more important, because this allow the developing software faithful to the algorithm language.
Finally, this paper show some relevant techniques used in models and to describe several details.
\end{abstract}
%
\section{Introduction about hardware models in B}
[One page]
This paper show a model in B from a microcontrooller.
[EXPLAIN THE MOTIVATION]
 [When a error is found at project of plataform the PREJUIZOS are big ]
 [ In critical system, when a error apper in software it can make several PREJUIZOS  ]

[[Other microcontrollers are in building proccess: 8051, Arm..]]
[Some smal details of the model are omitted in order to simplify the explanation.]
[Smal intrudctio about B]
[Structure from paper]

@@

\subsubsection{ Desing components B }
[half page]
[Explicar como as coias são modeladas]

\section{General definitions for hardware models}

@@

\begin{itemize}
  \item ALU \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and returns its decrement. There is a
  similar function named $\mathit{dec}$.

  \item ALU \it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the pc actual
  value and return its increment.

  \item ALU \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
  byte is zero, otherwise returns 0.

  \item byte \it is\_zero $\in$ BYTE  $\fun$   BIT \rm - It returns 1 if the received byte is zero, otherwise
  returns 0.

  \item byte \it rotateright  $\in$  \it BYTE  $\fun$  \it BYTE \rm - It receives a byte and returns its rotated to
  right. There is a similar function named $\mathit{rotateleft}$.

  \item ALU \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.
\end{itemize}
Every logic functions defined in the $\mathit{BYTE}$ module are included in the $\mathit{ALU}$ module. Then, it was
not necessary rewrite these functions.

\subsection{Interrupts}

@@

The Z80 has an extensive set of input and output instruction and 256 ports for
devices. This model can transfer data blocks and between the I/O devices and any
of the internal registers or memory address.

The $\mathit{IN\_r(C)}$ \footnote{The tools B don't allow use parenteses in identifiers, then the characteres
\{``('',``)''\} are replaced respectively by \{``9'',``0''\} in the real specification.} is represented by the follow B
operation. This instruction is from I/O group and update the flag registers.

\hspace*{0.0in}\bf IN\_r\_(C) \rm ( \it rr \rm ) \rm =

@@

\subsubsection{Instruction set} 
% Talk in general and put a example with midlle
% complex

@@

@@

@@

\bibitem {Dantas}
 Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Proposta e avaliação de uma abordagem de desenvolvimento de
 software fidedigno por construção com o método B. In: SEMISH, Belém, 2008. SBC.

>>>>

@@

\begin{abstract}
This paper describes an approach to model the microcontroller platforms. More specifically, it shows details about
the Z80 model. The model is created using the B method; it applies math and logic concepts to describe the
characteristics from platforms. Therefore, this model can be used in platforms projects to document, build
simulators, verify properties about the model and verify the software in assembly level. The verification in
assembly is utility more important, because this allow the developing software faithful to the algorithm language.
Finally, this paper show some relevant techniques used in models and to describe several details.
\end{abstract}
%
\section{Introduction}
% Explica o contexto geral do artigo
%  Porque modelar os microcontroladores.
%    Porque podemos verificar várias propriedades em tempo de projeto, documentar, simulacao através das ferramentas de suporte do método B...
%    Porque permite verificar software em nível de assemblagem.[citar trabalho de BHDL]
%  Contextualizacao com grande desafio da verificao formal em nivel de assemblagem[citar grande challenge][citar proposta do ALEMAO do ProB] 
%  Conclusão da introducao.
%  Organizacao do paper

%  Basear-se no Relatorio técnico e no ERMAC

@@

\subsubsection{Desing components B }
[half page]
[Explicar como as coias são modeladas]
% Use the tech report
\section{General definitions for hardware models}

@@

\begin{itemize}
  \item \it inc  $\in$ \it BYTE $\fun$  \it BYTE \rm - It receives a byte and returns its decrement. There is a
  similar function named $\mathit{dec}$.

  \item \it instruction\_next  $\in$  USHORT  $\fun$  USHORT \rm - It receives the pc actual
  value and return its increment.

  \item \it is\_negative  $\in$  \it BYTE  $\fun$  \it BIT \rm - It returns 1 if the received
  byte is zero, otherwise returns 0.

  \item \it update\_refresh\_reg \rm - It receives a byte and returns its increment until the seventh bit.

\end{itemize}

Every logic functions defined in the $\mathit{BYTE}$ and $\mathit{BV16}$ module are included in the $\mathit{ALU}$
module. Then, it was not necessary rewrite these functions.

\subsubsection{Modelling the instructions} 
[[Show and  LD n A]]

\subsection{Modelling the Interrupts}

@@

The Z80 has an extensive set of input and output (I/O) instruction and 256 ports for
devices. This model can transfer data blocks and between the I/O devices and any
of the internal registers or memory address.

The $\mathit{IN\_r(C)}$ \footnote{The tools B don't allow use parenteses in identifiers, then the characteres
\{``('',``)''\} are replaced respectively by \{``9'',``0''\} in the real specification.} instruction is represented
by the follow B operation. This instruction is from I/O group and update the flag registers.

\hspace*{0.0in}\bf IN\_r\_9C0 \rm ( \it rr \rm ) \rm =

@@

@@

 [Basear-se no ERMAC]

@@

\bibitem {CLEARSY}
CLEARSY. Atelier B. http://www.atelierb.eu.

\bibitem {Dantas_Semish08}
 Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Proposta e avalia\c{c}\~{a}o de uma abordagem de desenvolvimento de
 software fidedigno por constru\c{c}\~{a}o com o m\'{e}todo B. In: SEMISH, Bel\'{e}m, 2008. SBC.

\bibitem {Dantas_Report08}
Dantas, B; D\'{e}harbe, D; Galv\~{a}o, S; ET AL.Verified compilation based on the B method: initial
appraisal.Technical Report, UFRN - DIMAP, 2008.

@@

\bibitem {HOARE}
HOARE, C. A. R. The verifying compiler, a grand challenge for computing research. In: VMCAI, p. 78-78, 2005.

\bibitem{Medeiros}
 Val\'{e}rio Medeiros Jr., Stephenson Galv\~{a}o, David D\'{e}harbe. D\'{e}harbe, D; Galv\~{a}o, S; ET AL. Modelagem
 de Microcontroladores em B. In: ERMAC, Natal, 2008. SBMAC.

\bibitem {WORDSWORTH}
WORDSWORTH, J. B. Software Engineering with B. Addison Wesley, Boston, 1996.
